---
title: "Using the gPIPE Package for Dose-Finding Studies"
author: "Xiaoran Lai"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the gPIPE Package for Dose-Finding Studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 5
)
```


# Introduction
This vignette demonstrates the use of the gPIPE package for dose-finding studies. 
The package implements a generalized Product of Independent Beta Probabilities Escalation (PIPE) design with epsilon-tapering. The key components include:

- `Drug`: Manages drug information and dose levels
- `DrugCombination`: Handles combinations of drugs and their dose levels
- `DoseConfiguration`: Manages dose configurations
- `PipeEstimator`: Implements the core PIPE methodology
- `PatientDataModel`: Tracks patient data throughout the trial

# Getting Started
First, we load the required packages:
```{r pkg-setup, echo = TRUE}
library(gPIPE)
library(dplyr)
```

# Constructors

## Drug Class
The Drug class is designed to manage drug information, including its name and dose levels. 
It offers a versatile way of initializing drug data either programmatically or with parameters.

We can create the object directly:
```{r, echo = TRUE}
# Create a drug with 3 dose levels
num_doses <- 3
drugA <- createDrug(name = 'drugA', doseCount = num_doses)

# Create drugs with specific dose levels and labels
drug1 <- createDrug(name = 'Drug1', doseCount = num_doses)
drug2 <- createDrug(name = 'Drug2', doseCount = num_doses)

# Create a list of drugs for combination use
drug_list <- list(drug1, drug2)
```

### Working with Drug Objects

### Get dose informations
```{r, echo = TRUE}
# Get dose labels for Drug1
drug_labels <- drug1$getDoseLabels()
print(drug_labels)

# Get numeric level for "dose2" dose level
dose_levelA <- drugA$getDoseLevels("dose2")
print(dose_levelA)

# Add a new dose level
drugA$addDose(new_label = "Extra", new_level = 4)

# Get updated dose labels 
updated_labels <- drugA$getDoseLabels()
print(updated_labels)
```

## Drug Combination Class

The `DrugCombi` class, short for Drug Combination, is designed for managing combinations of different drugs and their respective dose levels. 
It is a composite constructor of multiple drugs that gives additional methods uniquely defined for drug combinations.

```{r, echo = TRUE}
# Create a drug combination object using our list of drugs
drugcombi <- createDrugCombi(drugs = drug_list)

# Get all possible dose combinations
dose_combinations <- drugcombi$getDoseCombinationsLevel()
print(dose_combinations)
```

## Dose Configuration Class

The DoseConfiguration class manages dose configurations for dose-finding studies. 
Let's create valid dose configurations that satisfy monotonicity constraints:

```{r dose-configs-best, echo = TRUE, cache=TRUE}
# Generate monotonic matrices for a dose grid
num_doses_a <- 3
num_doses_b <- 3
all_configs <- monotonic_matrices(num_doses_a, num_doses_b, "decreasing")
all_configs <- lapply(all_configs, function(x) c(x))

# Create dose configuration objects
dose_configs <- lapply(all_configs,
    function(x) createDoseConfig(
        drugCombiObject = drugcombi, 
        isNonDecreasing = FALSE, 
        currentConfig = as.numeric(x)
    )
)

# Keep only valid configurations
dose_configs_valid <- Filter(function(x) x$isValid, dose_configs)
```

## PIPE Estimator Class
The PipeEstimator class implements the core PIPE methodology. It takes a list of valid dose configurations and evaluates the posterior gain of each configuration.
The tuning parameters `epsilon` and `weight` can be fixed or adaptive at each update.

```{r, echo = TRUE}
# Create a PIPE estimator with linear tapering
pipe_hat <- createPipeEstimator(
    validConfigs = dose_configs_valid, 
    epsilonTarget = 0.5, 
    taper_type = 'linear'
)
```

## PatientDataModel Class

The PatientDataModel class tracks patient data throughout a trial:

Create a PatientDataModel object:
```{r, echo = TRUE}
# Create a patient data model
patientDataModel <- createPatientDataModel(drugcombi, startingDoseLevel = 'dose1.dose1')
patientDataModel$resetTrial()

# Add individual patient data
patientDataModel$addPatientData(doseCombination = "dose1.dose1", outcome = 0, cohort = 1)
patientDataModel$addPatientData(doseCombination = "dose2.dose2", outcome = 1, cohort = 2)
patientDataModel$addPatientData(doseCombination = "dose2.dose1", outcome = 1, cohort = 3)
```

Summary statistics for each dose level combination can be printed for furhter processing.
We can toggle between printing all dose levels or only those onto which patients enrolled.

```{r}
# Get summary statistics for doses that have patients enrolled
summaryStats <- patientDataModel$getSummaryStats(includeAllCombi = FALSE)
print(summaryStats)

# Generate random patient data up to a specified dose level
patientDataModel$generateRandomPatientData(
    doseLevel = "dose2.dose2",
    numPatients = 5,
    outcomeProb = 0.1
)

# View patient data
head(patientDataModel$patientData)
```

## Drug Combination Model Class
The `DrugCombinationModel` constructor allows the initialisation and update of the statistical model used for each
drug combination level. Depending on the model (currently only binomial and normal distribution for binary and continuous outcomes, respectively)
, it reads in prior information, including hyperparameters, from a YAML file, and update the posterior of interest based on the patient data that are
fed into the model.
```{r, echo = TRUE}
# Create a DrugCombinationModel object with default priors
config_list <- list(
    theta = 0.3,
    priorA = 0.5,
    priorB = 0.5
)

drugCombinationModel <- createDrugCombinationModel(
    config_list = config_list,
    family = "binomial"
)

# Update the model based on patient data
updatedModel_posterior <- drugCombinationModel$updateModel(patientDataModel)

# View a sample of the updated posterior probabilities
head(updatedModel_posterior)
```

# PIPE estimator update

Once the posterior for the parameter of interest is calculated, we can then pass it onto the `PipeEstimator` to update
the posterior gain to select the best config so far
```{r, echo = TRUE}
# Update PIPE estimator with posterior probabilities
pipe_hat$updatePipeEstimator(updatedModel_posterior)
```

# Conducting a dual-agent dose-escalation trial using PIPE design
## Initialisation
Let us first define the dose levels for each drug under investigation in the trial. 
This can be done either through reading information from a YAML file with specific dose levels, or initialised without.

```{r, echo = TRUE}
# Set up drugs for the trial
num_doses_a <- 5
num_doses_b <- 3
drugA <- createDrug(name = 'drugA', doseCount = num_doses_a)
drugB <- createDrug(name = 'drugB', doseCount = num_doses_b)
drug_list <- list(drugA, drugB)

# Create drug combination
drugcombi_new <- createDrugCombi(drugs = drug_list) 

# Get available dose combinations
dose_combinations <- drugcombi_new$getDoseCombinationsLevel()
names(dose_combinations)[1:5]
```

We will then set up the PIPE configurations and find the set of configurations that fulfills the partial-order constraint.
Note that we would like our dose configuration to be equal to 1 if it the dose level is acceptable, and 0 otherwise.
This means that we will set `isNonDecreasing = FALSE` for non-increasing configurations.
```{r new, echo = FALSE, cache=FALSE}
# Generate valid dose configurations
all_configs <- monotonic_matrices(num_doses_a, num_doses_b, "decreasing")
all_configs <- lapply(all_configs, function(x) c(x))
dose_configs_new <- lapply(all_configs,
    function(x) createDoseConfig(
        drugCombiObject = drugcombi_new, 
        isNonDecreasing = FALSE, 
        currentConfig = as.numeric(x)
    )
)
dose_configs_valid <- Filter(function(x) x$isValid, dose_configs_new)

# Create PIPE estimator
pipe_hat <- createPipeEstimator(
    validConfigs = dose_configs_valid,
    epsilonTarget = 0.5, 
    taper_type = 'linear'
)

# Set up admissibility rules and selection strategies
# Note: These classes are available in the package

# Set trial parameters
current_dose_level <- "dose1.dose1"
cohort_size <- 2
max_cohorts <- 20
```

# Conclusion
The gPIPE package provides a comprehensive framework for implementing the generalized PIPE design with epsilon-tapering for dose-finding studies. This vignette has demonstrated the key components and workflow for designing and analyzing such studies.

# References

1. Mander, A. P., & Sweeting, M. J. (2015). A product of independent beta probabilities dose escalation design for dual-agent phase I trials. *Statistics in medicine*, 34(8), 1261-1276. <doi:10.1002/sim.6434>

2. Cheung, Y. K., Chandereng, T., & Diaz, K. M. (2022). A novel framework to estimate multidimensional minimum effective doses using asymmetric posterior gain and epsilon-tapering. *The annals of applied statistics*, 16(3), 1445. <doi:10.1214/21-AOAS1603>
